Different forms of uncertainty differently affect the evolution of social learning

### 2022-08-19 ms.tex
Updates from Overleaf

\begin{abstract} 
  Social learning is essential to human survival---it is a critical
  adaptation for dealing with uncertainty. However, uncertainty takes many forms.
  We identified four theoretically important types of uncertainty: temporal
  environmental variability, ambiguous payoffs, decision set size, and effective
  lifespans. We used this framework to study conditions for the evolution of
  social learning in populations with cognitively realistic individual learning
  capacities.  We used evolutionary agent-based modeling to track how often social, rather than individual,
  learning evolved. Agents performed one of several behaviors to acquire payoffs.
  Individual learning used the softmax algorithm. Social learning occurs between
  generations  through vertical and oblique transmission.  Different types of
  uncertainty had different effects. Temporal environmental variability suppressed
  social learning. Larger decision set sizes promoted social learning even
  when the environment changed frequently. Payoff
  ambiguity and lifespan had more complex effects on social learning, dependent on
  other uncertainty parameters.  This shows the importance of clearly
  operationalizing uncertainty in theories of social learning evolution.
% Social learning is essential to human survival. It is likely to evolve when it is more
% efficient than asocial, trial-and-error learning. Theoretically, social learning
% is adaptive for some uncertainty, but too much uncertainty makes
% social information unreliable. This fact is empirically
% supported across biology and human sciences. However, it is unclear how
% specific classes and types of uncertainty affect social
% learning evolution. Furthermore, existing models and experimental operationalizations of
% uncertainty are ambiguously related, and 
% tend to only consider a small number of behavioral
% choices.  Here we use evolutionary agent-based modeling to consolidate
% models of uncertainty in social learning evolution to address these issues.
% We model societies of agents who perform one of a number
% of possible behaviors to acquire payoffs in a time-varying environment.
% We identified complex patterns in social learning evolution depending on contextual
% uncertainty variables: environmental variability, ambiguous payoffs, 
% decision sets size, and effective lifespans. 
% Our work advances social learning evolution theory in a way that could help 
% guide human, non-human, and artificial intelligence towards optimal responses 
% to existential threats and new opportunities.
\footnote{This document contains
\total{citnum} references.}  
\end{abstract}

@@

Social learning enhances problem solving when acquiring information from others is more
efficient than learning on one's own~\cite{Laland2004}. However, social learning can also
lead individuals astray if, for example, they are copying outdated information.
Theoretical models have helped clarify the circumstances under which social learning
should be evolutionarily favored~\cite{BoydRicherson1985}. This theoretical work has
motivated much empirical research and helped predict social learning behavior in humans
and other species~\cite{McElreath2005,Kendal2018,Allen2019}.  This literature
demonstrates that social learning is a critical adaptation found across taxa for dealing
with uncertainty.  Uncertainty weighs particularly heavily on human adaptation because of
our long lifespans, cosmopolitan distribution, and the resulting necessity to adapt to
relatively coarse-grain environments \cite{levins1962}. However, the term ``uncertainty"
often conflates several different interpretations of that word. For example, different
models define uncertainty in terms of the rate of environmental change, spatial
environmental variation, or the reliability of information acquired from the environment.
Furthermore, in translating mathematical models to verbal predictions, the formal meaning
of terms like uncertainty is often lost \cite{lawson1988probability}, facilitating such
conflation. Furthermore, it can be important to distinguish a narrower definition of
uncertainty wherein the probabilities of outcomes remain unknown~\cite{knight1921,
volz2012}. As the number of formal models of social learning have expanded, an ever
increasing number of modeling choices~\cite[Figure 1]{Kendal2018} and formalizations of
uncertainty have made it difficult to compare across models and consolidate our
understanding of the contexts in which social learning is adaptive. 
% Finally, while risk and uncertainty are often used interchangeably to refer to
% non-deterministic outcomes,

Other (asocial) forms of learning are also likely adaptations to uncertainty that should be
accounted for, but are typically glossed over in models.  In general, learning is an iterative
process by which one repeatedly acquires information, forms representations and
predictions, and tests and refines those representations and predictions to manage
uncertainty~\cite{jacobs2011bayesian,clark2013whatever}.  Because cultural evolutionary
models are more often used to explain social learning mechanisms, they tend to use a
single, weak individual-learning mechanism for comparison. Such modelling choices may lead us to underestimate the extent to which individual learning provides
quality information, both for asocially adapting to uncertainty and for social learners to take advantage of.
%and for the individual to adapt to the challenges
assumed to be solved by social learning. 
% underestimate the quality of information
% available for social learning.

To address these concerns we developed an evolutionary agent-based model that
simultaneously operationalizes several forms of uncertainty and endows agents
with a relatively powerful learning mechanism. We model four kinds of
uncertainty that are common in cultural evolution models and the related empirical
literature:  (1) temporal environmental variability; (2) selection set size (the number of
possible behaviors); (3) payoff ambiguity (the difference in the expected rewards for
different behaviors); and (4) the effective lifespan of agents (the number of
opportunities for individual learning), explained below. The model also assumes a more
cognitively realistic and empirically-motivated learning mechanism, the softmax algorithm.
This allows for a more accurate assessment of tradeoffs between 
social and asocial learning
by fleshing out an individual-level learning mechanism that makes full use of 
the individually and socially learned data. 
In this work, social learners learn from the previous generation, and 
asocial learners do not. All agents engage in individual-level learning within
their lifespans whether they are social learners or not.
We use this integrated model to ask which kinds of uncertainty are likely to
favor social learning and to clarify the logic underlying the results.
%of why this would be the case.

In the remainder of this introduction we describe previous research on the forms of uncertainty and the learning mechanism we include in our model. We then present how our evolutionary agent-based model works. Finally, we explain the logic underlying the result that different forms of uncertainty can either suppress or encourage the evolution of social learning.

@@

\textbf{Temporal environmental variability.} When environments change
(e.g., climatic events, migration, technological paradigm shifts),
strategies that were previously adaptive may no longer be optimal. The difficulty
in predicting either when such shocks will occur or which behaviors will be
adaptive in the resulting environments leads to uncertainty.

Temporal environmental variability is fundamental to most evolutionary models of
social learning. Such fluctuations have been proposed as an important selective
pressure for learned as opposed to genetically fixed behaviors
~\cite{Richerson2000}. On the other hand, environments that change too quickly
will lead the population away from social learning to avoid outdated, 
and therefore maladaptive information~\cite{Feldman1996,
BoydRicherson1985}. This suggests that intermediate levels of temporal variation
are important for the evolution of social learning mechanisms~\cite{aoki2005}.

Despite the theoretical centrality of temporal environmental variability, this construct is
implemented in several ways~\cite{aoki2014evolution}. Most commonly, it is modeled
as an independent probability of the environment changing its state each
generation
~\cite{BoydRicherson1985,Rogers1988,Feldman1996,McElreath2005,Enquist2007,perreault2012bayesian,aoki2014evolution}.
% The probabilities of environmental change are usually fixed per generation
%~\cite{BoydRicherson1985,Rogers1988,kendal2009evolution}, 
but has also been modeled as deterministic cycles~\cite{Feldman1996, aoki2014evolution}.
Furthermore, the consequences of environmental change can range from mild to
catastrophic. In the latter case, a change of environment results in the total
elimination of any adaptive behavior, which must be learned \emph{de novo} for the
new environment~\cite{Rogers1988}. Such catastrophic environmental changes present
a large adaptive challenge as individuals cannot rely on accumulated information
from previous generations. Other models of environmental change introduce less
uncertainty as they fluctuate between two environmental states with corresponding
adaptive behaviors. This means that previously maladaptive
behaviors become adaptive when the environment changes
~\cite{perreault2012bayesian}. The chosen mechanism for modelling temporal
change has real theoretical consequences~\cite{aoki2014evolution}.

\textbf{Selection set size.}
When one does not know which option to take, uncertainty increases with the number of options, which we call the selection set size.
% This is a mathematical fact at the
% core of information theory: the number of bits required to reliably encode the optimal solution increases by one every time the
% selection set doubles in size~\cite{mackay2003information}. 
% This is intuitive; the more options to consider, the greater one's uncertainty about which option is the best choice. 
In studies of social learning, the selection set is often limited
to just two options.  For example, in empirical studies of \emph{bombus
terresteris} (bumble bee)~\cite{Baracchi2018} and \emph{parus major} (great
tit)~\cite{Aplin2017}, experimenters provided two behaviors for the bees or birds
to choose from, where one yielded a higher payoff than the other. %respectively,
each yielding greater or smaller payoffs depending on experimental treatment.
Similarly, many human studies have used just two or three possible
behaviors~\cite{McElreath2005,Toyokawa2019}. Behavioral study designs are
often motivated by modeling studies with similar formulations~\cite{Rogers1988,boyd1995does,
perreault2012bayesian}.
% \cm{pretty sure boyd1995 uses same trick as
% rogers 88 so not sure why they were being used as illustrating separate kinds of
% behavior selection sets. correct if wrong} 
Larger, more open sets of behavioral
choices are not uncommon in experiments trying to capture more complex or
naturalistic tasks~\cite{derex2013, wasielewski2014}, but these map on
imperfectly to the theoretical literature that tend to use a limited set of behavioral options.
%\mt{How so?}. 
Some models have studied systems with
larger, but defined, selection sets~\cite{Rendell2010}, while in others the
number of options is implicit in the probability that a learner gets the right answer~\cite{Enquist2007}.
%\mt{How does this work exactly in Enquist et al 2007?}. 
Researchers have also compared models with a different number
of environment states, corresponding to a different number of optimal behaviors
through time~\cite{Feldman1996}. These examples share one common feature: 
the size of the selection set was held static \emph{within} model generations.
% as a measure of changing uncertainty in cultural evolutionary models.

\textbf{Payoff ambiguity.} Many models of social learning differentiate between
the payoffs for adopting optimal vs. non-optimal behaviors
\cite{Rogers1988,Enquist2007,Rendell2010}. The size of the difference between
these payoffs is usually taken to indicate the strength of selection of learning,
which makes sense. However, in reality payoffs for particular behaviors are not
always consistent. A behavior may yield a large payoff sometimes and a small
payoff other times \cite{McElreath2005}. This means that signals about the
relationship between behavior and payoff are often noisy, and differentiating
between behavioral options is in part a problem of signal detection.  When the
difference in expected payoffs between optimal and non-optimal behaviors is very
large, this noise matters little, as the signal is still very clear. However, when
the expected payoffs of different behaviors are similar relative to the size of
their variances, ambiguity arises about which behaviors really are superior.
Smaller differences between payoffs corresponds to larger ambiguity. Payoff
ambiguity has been manipulated in both theoretical~\cite{perreault2012bayesian}
and empirical~\cite{McElreath2005, Morgan2012} studies, both of which support the
claim that payoff ambiguity increases the reliance on social information.
Importantly, payoff ambiguity %differs from selection set size in that it not only
affects uncertainty, it also directly influences both uncertainty \emph{and} the
strength of selection. %This parameter sometimes takes the form of differences in
expected payoffs arising from different behaviors or
strategies~\cite{Enquist2007,Rendell2010} or by varying the standard deviation of
payoffs~\cite{McElreath2005}.

\textbf{Effective lifespan.} 
The more opportunities one has for
learning in their lifespan, the more uncertainty can be reduced, assuming a
static environment. Similarly, a reduction in the number
of opportunities to learn will increase the uncertainty about whether one has
accurately learned about the available behavioral options and their associated
payoffs. That uncertainty is passed on to future generations.
% Learning is an iterative process by which one repeatedly acquires information, forms
% representations and predictions, and tests and refines those representations and
% predictions to manage uncertainty \cite{jacobs2011bayesian,clark2013whatever}. 
We refer to the number of learning opportunities as an individual's
effective lifespan to imply that it is the number of opportunities to learn
throughout the individual's lifespan that matters here, not necessarily how long
they live in the absence of relevant learning opportunities.

Empirically, the number of learning opportunities can be manipulated in the lab,
but will also correspond with an individual's age the real world. In multi-round
studies of information use in novel tasks, US participants' use of social
information declined precipitously across rounds~\cite{McElreath2005}, suggesting
they were more likely to use social information when they were most uncertain
about the task early on. In a more naturalistic context, Aplin et al. (2017)
\nocite{Aplin2017} found that younger great tits more readily used social
information compared with older individuals, possibly because they had accrued
less information via individual learning, and also possibly because younger
individuals have the most to gain (given their expected remaining lifespan) by
switching to superior behavioral options. 
% \ps{I got rid of the Leris and Reader reference because it didn't seem relevant
% to age-based uncertainty. They just showed that young guppies exposed to
% reliable models were more likely to use social information later. I also added
% some human studies, which seems important.} 
Cross-cultural studies on humans have shown the importance of childhood as a phase
of heavy social learning~\cite{Reyes2016}. Young children are more likely to
acquire their beliefs and simple skills from their parents than are older children
or adults, which is at least partly due to the differential knowledge accrual
between young children and the older adults to whom they direct the most trust and
attention \cite{kline2013teaching}.
%The age of individuals was found to have an effect on social learning in both
%\emph{parus major}~\cite{Aplin2017} and \emph{poecilia reticulata}
%(guppies)~\cite{Leris2016}. \ps{What effect? I'm not sure } \mt{Need better
%connection between lifespan and age, or better references to support use of
%lifespan}.
The effective lifespan for learning
varies across tasks, individuals, and species, yet most models assume one learning
opportunity per generation~\cite{Feldman1996,Henrich1998, perreault2012bayesian}.
Some models do allow several cultural generations within 1 genetic generation
~\cite{Enquist2007}, but little formal theory explicitly examines the role of
learning opportunities in social learning evolution models.

@@

Social learning is the adaptive icing on the cake of other cognitive adaptations
to ecological uncertainty.  While the literature reviewed above suggests that
several forms of uncertainty play a role in favoring the evolution of social
learning, several other cognitive mechanisms have also evolved in response to
uncertainty~\cite{volz2012}. Many of these are flexible learning mechanisms that
do not require assessing others' information. For example, people in experimental
studies increase their exploratory behavior when total uncertainty in the task
increases and are more likely to test %Furthermore, correlations in uncertainty
between potential behaviors, humans tend to adopt directed exploration strategies
that favor testing behaviors with greater observed payoff variance
\cite{Wilson2014,Gershman2019}.

Many models of social learning assume little to no cognitive capabilities
of their agents. For example, a common modelling strategy compares the payoffs of
agents with different pure learning strategies (some will always engage in social
learning, and others in individual learning) while revealing nothing about the
cognition underlying individuals' decisions~\cite{BoydRicherson1985, Rogers1988,
aoki2005}. Some more complicated learning strategies that make use of both
socially and individually learned information have been proposed
~\cite{Enquist2007, perreault2012bayesian}, but these are seldom designed to
reflect cognitive mechanisms for integrating information. Though such
simplifications are often useful, several authors have warned of the risks to
\emph{blackboxing}, or ignoring, cognitive mechanisms \cite[p. 658]{Heyes2016,
Kendal2018}, given such details may be theoretically meaningful. In order to
understand how social learning evolves as an adaptation to uncertainty, we 
endowed agents with a biologically plausible cognition capable of integrating
information from various sources to maximize payoffs, \emph{softmax
search}~\cite{Gershman2019}, described in more detail below. 
% \cm{I don't
%   understand this next sentence, but think it can be cut with no loss of meaning:
%   In our model we do not impose any structure in randomness between behaviors, so
% our cognitive model is only sensitive to total randomness.} 
% To do so, we let
% agents see the payoffs to behaviors and use the softmax function to convert these
% observations to a probability distribution from which they strategically select
% their own behavior. %\cm{does this still make sense? and does it need a citation?}

@@

To understand how uncertainty affects social learning, 
we developed an agent-based model of
social learning evolution.  Evolutionary agent-based models simulate changes in
the prevalence of a phenotype over several generations---in this model the phenotype
is whether or not an agent engages in social learning.  In our model, a
population of $N$ individuals each must decide which of $B$ behaviors to perform at
each time step within a generation consisting of $L$ time steps. At the end of each
generation, agents are selected with replacement to reproduce another $N$ agents, 
biased by net payoffs. Agents
who inherit the social learning trait then learn about payoffs from the
previous generation, while asocial learners begin life with no \emph{a priori}
knowledge of the world.  Each behavior is a ``bandit'', a common modeling and
experimental approach for representing behaviors with random-valued
payoffs~\cite{SuttonBartoBook,McElreath2005,Steyvers2009,Rendell2010,Schulz2019}.
Bandits in our model yield Bernoulli payoffs of 1 or 0 with some probability,
with that probability defining each behavior's expected payoffs.  One of the
bandits/behaviors in the model is more likely to pay off than all the rest.  Agents
optimize their net payoffs over their lifespan by quickly figuring out which
behavior is optimal and performing it as often as possible within their lifespan.
The optimal behavior is static within agent lifespans, but varies in general between
generations, representing environmental variability.

>>>>

\begin{abstract} 
  Social learning is %essential to human survival---it 
  is a critical
  adaptation for dealing with uncertainty. However, uncertainty takes many forms.
  We identified four distinct and theoretically important types of uncertainty: temporal
  environmental variability, ambiguous payoffs, decision set size, and effective
  lifespans. We embed this framework in an evolutionary agent-based model to study conditions for the evolution of  social learning in populations with cognitively realistic individual learning
  capacities. % We used evolutionary agent-based modeling to track the evolution of social learning versus purely individual learning.
  %how often social, rather than individual,   learning evolved. 
  Agents could perform one of several behaviors, modeled as multi-armed bandit, to acquire payoffs.
  Individual learning used the softmax algorithm.  Social learning involved vertical and oblique transmission between generations, which served as a constraint on subsequent individual learning.  We found that different types of
  uncertainty had different effects. Temporal environmental variability suppressed
  social learning. Larger decision set sizes promoted social learning even
  when the environment changed frequently. Payoff
  ambiguity and lifespan had more complex effects on social learning, dependent on
  other uncertainty parameters.  This study shows the importance of clearly
  operationalizing uncertainty for understanding its role in social and cultural evolution.
% Social learning is essential to human survival. It is likely to evolve when it is more
% efficient than asocial, trial-and-error learning. Theoretically, social learning
% is adaptive for some uncertainty, but too much uncertainty makes
% social information unreliable. This fact is empirically
% supported across biology and human sciences. However, it is unclear how
% specific classes and types of uncertainty affect social
% learning evolution. Furthermore, existing models and experimental operationalizations of
% uncertainty are ambiguously related, and 
% tend to only consider a small number of behavioral
% choices.  Here we use evolutionary agent-based modeling to consolidate
% models of uncertainty in social learning evolution to address these issues.
% We model societies of agents who perform one of a number
% of possible behaviors to acquire payoffs in a time-varying environment.
% We identified complex patterns in social learning evolution depending on contextual
% uncertainty variables: environmental variability, ambiguous payoffs, 
% decision sets size, and effective lifespans. 
% Our work advances social learning evolution theory in a way that could help 
% guide human, non-human, and artificial intelligence towards optimal responses 
% to existential threats and new opportunities.
\footnote{This document contains
\total{citnum} references.}  
\end{abstract}

@@

Social learning enhances problem solving when acquiring information from others is more
efficient than learning on one's own~\cite{Laland2004}. However, social learning can also
lead individuals astray if they are copying irrelevant or outdated information.
Theoretical models have helped clarify the circumstances under which social learning
should be evolutionarily favored~\cite{BoydRicherson1985, aoki2014evolution,Kendal2018}, and these models have motivated empirical work that has both validated and refined theoretical predictions concerning social learning behavior in humans
and other species~\cite{galef2005social,McElreath2005,Kendal2018,Allen2019}.  This literature
indicates that social learning is a critical adaptation found across taxa for dealing
with uncertainty. When in doubt, copy others.

Uncertainty weighs particularly heavily on human adaptation because of
our long lifespans, cosmopolitan societies, and dispersal across a highly variable environments, requiring coarse-grain and plastic behavioral adaptations \cite{levins1962}. However, the term ``uncertainty"
often conflates several different interpretations of that word. For example, different
models define uncertainty in terms of the rate of environmental change, spatial
environmental variation, or the reliability of information acquired from the environment.
Furthermore, in translating mathematical models to verbal predictions, the formal meaning
of terms like uncertainty is often lost \cite{knight1921, lawson1988probability, volz2012}, facilitating such
conflation. %Furthermore, it can be important to distinguish a narrower definition of uncertainty wherein the probabilities of outcomes remain unknown~\cite{knight1921,volz2012}. 
As the number of formal models of social learning have expanded, an ever
increasing number of modeling choices~\cite{Kendal2018} and formalizations of
uncertainty have made it difficult to compare across models or to consolidate our
understanding of the contexts in which social learning under uncertainty is adaptive. 
% Finally, while risk and uncertainty are often used interchangeably to refer to
% non-deterministic outcomes,

%Other (asocial) forms of learning are also likely adaptations to uncertainty that should be accounted for, but are typically glossed over in models.  
In general, learning is an iterative
process during which an individual acquires information, forms representations and
predictions, and then tests and refines those representations and predictions to manage
uncertainty~\cite{jacobs2011bayesian,clark2013whatever}.  Because cultural evolutionary
models are often designed to explain \emph{social} learning mechanisms, they often contrast social learning with an overly simplistic mechanism for individual learning. In particular, it is often assumed that individual learners can \emph{always} pay a cost to learn the optimal behavior for a particular environment with certainty. This assumption implies that social learning by observing an individual learner is a good bet if the environment has not changed. Such modelling choices may lead us to overestimate both the extent to which individual learning provides
quality information and the value of social learning in a population of individual learners. % , both for asocially adapting to uncertainty and for social learners to take advantage of.
%and for the individual to adapt to the challenges assumed to be solved by social learning. 
% underestimate the quality of information
% available for social learning.

To address these concerns we developed an evolutionary agent-based model that
simultaneously operationalizes several forms of uncertainty and endows agents
with a relatively powerful mechanism for individul learning. We model four kinds of
uncertainty that are common in cultural evolution models and the related empirical
literature:  (1) temporal environmental variability; (2) selection set size (the number of
possible behaviors); (3) payoff ambiguity (the difference in the expected rewards for
different behaviors); and (4) the effective lifespan of agents (the number of
opportunities for individual learning), explained below. In our model, social learners learn from the previous generation and asocial learners do not. All agents engage in individual learning within
their lifespans. The model also assumes a more
cognitively realistic and empirically-motivated learning mechanism, the softmax algorithm, which allows  learners to make full use of the available data acquired by both individual and social learning. 
Our model design thereby affords us a richer and possibly more accurate assessment of tradeoffs between 
social and asocial learning. 
We use this integrated model to ask which kinds of uncertainty are likely to favor social learning and to clarify the logic underlying the results.
%of why this would be the case.  
In the remainder of this introduction we briefly review previous research on both the forms of uncertainty and the learning mechanism used in our model. 
%We then present how our evolutionary agent-based model works. Finally, we explain the logic underlying the result that different forms of uncertainty can either suppress or encourage the evolution of social learning.

@@

\textbf{Temporal environmental variability.} When environments change
(for example, in response to climatic events, migration, or technological paradigm shifts),
strategies that were previously adaptive may no longer be optimal. The difficulty
in predicting either when such shocks will occur or which behaviors will be
adaptive in the resulting environments leads to uncertainty.

Temporal environmental variability is fundamental to most evolutionary models of
social learning. Such fluctuations have been proposed as an important selective
pressure for learned as opposed to genetically fixed behaviors when genetic adaptation cannot keep pace with environmental change 
~\cite{Richerson2000}. On the other hand, environments that change \emph{too} quickly
will select against social learning so that individuals avoid learning outdated--- 
and therefore maladaptive---information~\cite{Feldman1996,
BoydRicherson1985}. This suggests that intermediate levels of temporal variation
are important for the evolution of social learning mechanisms~\cite{aoki2005}.

%Despite the theoretical centrality of t
Temporal environmental variability tends to be modeled in one of 
several ways~\cite{aoki2014evolution}. Most commonly, it is modeled
as an independent probability that the environment changes its state (and its corresponding optimal behavior) before each new generation
~\cite{BoydRicherson1985,Rogers1988,Feldman1996,McElreath2005,Enquist2007,perreault2012bayesian,aoki2014evolution}, but
% The probabilities of environmental change are usually fixed per generation
%~\cite{BoydRicherson1985,Rogers1988,kendal2009evolution}, 
it has also been modeled using deterministic cycles, so that environments repeat at regular intervals ~\cite{Feldman1996, aoki2014evolution}.
%Furthermore, t
The consequences of environmental change can range from mild to
catastrophic. In the latter case, a change of environment results in the total
elimination of any adaptive behavior, which must be learned \emph{de novo} for the
new environment~\cite{Rogers1988}. Such catastrophic environmental changes present
a large adaptive challenge as individuals cannot rely on accumulated information
from previous generations. Other models of environmental change introduce less
uncertainty as they fluctuate between two or more environmental states with corresponding
adaptive behaviors. This means that previously maladaptive
behaviors become adaptive when the environment changes
~\cite{perreault2012bayesian}. The chosen mechanism for modelling temporal
change has important theoretical consequences, reviewed in Aoki \& Feldman (2014).~\nocite{aoki2014evolution} Our study design ignores cases where temporal variation is low enough that genetic selection can canalize a behavior, and instead considers only behaviors that can be learned.

\textbf{Selection set size.}
When one does not know which option to take, uncertainty increases with the number of options, which we call the selection set size.
% This is a mathematical fact at the
% core of information theory: the number of bits required to reliably encode the optimal solution increases by one every time the
% selection set doubles in size~\cite{mackay2003information}. 
% This is intuitive; the more options to consider, the greater one's uncertainty about which option is the best choice. 
In many studies of social learning, the selection set is often limited to just two options.  For example, in empirical studies of \emph{bombus
terresteris} (bumble bee)~\cite{Baracchi2018} and \emph{parus major} (great
tit)~\cite{Aplin2017}, experimenters provided two behaviors from which the bees or birds
could choose, with one yielding a higher payoff than the other. %respectively,
%each yielding greater or smaller payoffs depending on experimental treatment.
Similarly, many human studies have used just two or three possible
behaviors~\cite{McElreath2005,Morgan2012, Toyokawa2019}. Behavioral study designs are
often motivated by modeling studies with similar formulations~\cite{Rogers1988,boyd1995does,Feldman1996,
perreault2012bayesian}.
% \cm{pretty sure boyd1995 uses same trick as
% rogers 88 so not sure why they were being used as illustrating separate kinds of
% behavior selection sets. correct if wrong} 
Larger, more open sets of behavioral
choices are not uncommon in experiments trying to capture more complex or
naturalistic tasks~\cite{derex2013, wasielewski2014}, but these map on
imperfectly to the theoretical literature that tend to use a limited set of behavioral options.
%\mt{How so?}. 
Some models have studied systems with
larger, but defined, selection sets~\cite{Rendell2010, lindstrom2016co}, while in others the
number of options is subsumed under the probability that a learner gets the right answer~\cite{Feldman1996,Enquist2007}.
%\mt{How does this work exactly in Enquist et al 2007?}. 
%Researchers have also compared models with a different number of environment states, corresponding to a different number of optimal behaviors through time~\cite{Feldman1996}. 
Rarely if ever is the size of the selection set explored explicitly as a source of varying uncertainty, even though the number of options one has is likely to increase with the difficulty of the decision task~\cite{haynes2009testing,white2009testing}.  
%These examples share one common feature: the size of the selection set was held static \emph{within} model generations.
% as a measure of changing uncertainty in cultural evolutionary models.

\textbf{Payoff ambiguity.} Most models of social learning necessarily differentiate between the payoffs for adopting optimal vs. non-optimal behaviors
\cite{BoydRicherson1985,Rogers1988,Enquist2007,Rendell2010,aoki2014evolution}. The size of the difference between
these payoffs is usually taken to indicate the strength of selection of learning. %, which makes sense. 
However, the ability to discern payoff differences between behaviors is also a source of uncertainty. In reality, payoffs for particular behaviors are not always consistent. A behavior may yield a large payoff sometimes and a small
payoff other times \cite{McElreath2005}. This means that signals about the
relationship between behavior and payoff are often noisy, and differentiating
between behavioral options is in part a problem of signal detection.  When the
difference in expected payoffs between optimal and non-optimal behaviors is very
large, this noise matters little, as the signal is still very clear. However, when
the expected payoffs of different behaviors are similar relative to the size of
their variances, ambiguity arises about which behaviors really are superior.
Smaller differences between payoffs corresponds to larger ambiguity. Payoff
ambiguity has been manipulated in both theoretical~\cite{perreault2012bayesian}
and empirical~\cite{McElreath2005, Morgan2012} studies, both of which support the
claim that payoff ambiguity increases the reliance on social information.
Importantly, payoff ambiguity %differs from selection set size in that it not only 
affects both 
%affects uncertainty, it also directly influences both 
uncertainty \emph{and} the
strength of selection, with smaller payoff differences leading to greater uncertainty for a learner and weaker evolutionary selection favoring optimal strategies. %This parameter sometimes takes the form of differences in
%expected payoffs arising from different behaviors or strategies~\cite{Enquist2007,Rendell2010} or by varying the standard deviation of payoffs~\cite{McElreath2005}.

\textbf{Effective lifespan.} 
The more opportunities one has to learn during one's lifespan, the more uncertainty can be reduced, assuming a
static environment. Correspondingly, a reduction in the number of opportunities to learn will increase the uncertainty about which behavioral options are available and what their associated
payoffs are. %Such uncertainty can be passed on to future generations.
% Learning is an iterative process by which one repeatedly acquires information, forms
% representations and predictions, and tests and refines those representations and
% predictions to manage uncertainty \cite{jacobs2011bayesian,clark2013whatever}. 
We refer to the number of learning opportunities as an individual's effective lifespan to highlight that it is the number of opportunities to learn about the payoffs associated with a behavior, rather than the number of sunrises one experiences, that determines the level of uncertainty present in one's choices.  
%throughout the individual's lifespan that matters here, not necessarily how long they live in the absence of relevant learning opportunities.

Empirically, the number of learning opportunities can be manipulated in the lab, and in the real world will tend to correlate with an individual's relative age. In multi-round
studies of information use in novel tasks, US participants' use of social
information declined precipitously across rounds~\cite{McElreath2005}, suggesting
they were more likely to use social information when they were most uncertain about the task early on. In a more naturalistic context, Aplin et al. (2017)
\nocite{Aplin2017} found that younger great tits more readily used social
information compared to older individuals, possibly because they had accrued
less information via individual learning, and possibly because younger
individuals have the most to gain (given their expected remaining lifespan) by switching to superior behavioral options. 
% \ps{I got rid of the Leris and Reader reference because it didn't seem relevant
% to age-based uncertainty. They just showed that young guppies exposed to
% reliable models were more likely to use social information later. I also added
% some human studies, which seems important.} 
Cross-cultural studies have highlighted the importance of childhood as a phase
of heavy social learning in humans ~\cite{Reyes2016}. Young children are more likely to
acquire their beliefs and simple skills from their parents than are older children
or adults, which is at least partly due to the differential knowledge accrual
between young children and the older adults to whom they direct the most trust and
attention \cite{kline2013teaching}.
%The age of individuals was found to have an effect on social learning in both
%\emph{parus major}~\cite{Aplin2017} and \emph{poecilia reticulata}
%(guppies)~\cite{Leris2016}. \ps{What effect? I'm not sure } \mt{Need better
%connection between lifespan and age, or better references to support use of
%lifespan}.
The effective lifespan for learning
varies across tasks, individuals, and species, yet most models assume only one learning
opportunity per generation~\cite{BoydRicherson1985, Feldman1996,Henrich1998, perreault2012bayesian}.
Some models do allow several cultural generations within one genetic generation
~\cite{Enquist2007,Rendell2010,lindstrom2016co}, but little formal theory explicitly examines the role of
learning opportunities in the evolution of social learning.

@@

%Social learning is the adaptive icing on the cake of other cognitive adaptations to ecological uncertainty. 
While the literature reviewed above suggests that
several forms of uncertainty have played a role in the evolution of social
learning, other cognitive mechanisms have likely also evolved in response to
uncertainty~\cite{volz2012,johnson2013evolution,van2018uncertainty}. Many of these are flexible learning mechanisms that
do not require imitating the behaviors of others. For example, when faced with greater uncertainty individuals may adopt more exploratory learning strategies and may even preferentially test behaviors with greater observed payoff variance 
\cite{Wilson2014,Gershman2019}.

Many models of social learning simplify by using minimally cognitive %assume little to no cognitive capabilities of their 
agents. For example, a common modelling strategy compares the payoffs of
agents with different pure learning strategies (e.g., those who only engage in social
learning versus those who only engage in individual learning) while revealing nothing about the
cognition underlying individuals' decisions~\cite{BoydRicherson1985, Rogers1988,
aoki2005}. This sort of behavioral gambit is common in evolutionary modeling, but has also been criticized as `blackboxing' key cognitive processes that are important to social and cultural evolution \cite{Heyes2016,Kendal2018}.  
Some more complicated learning strategies that make use of both socially and individually learned information have been studied ~\cite{Enquist2007, perreault2012bayesian}, including those that integrate cognitively plausible mechanisms such as reinforcement learning \cite{lindstrom2016co}, though the latter approach remains the exception rather than the norm. 
%but these are seldom designed to reflect cognitive mechanisms for integrating information. Though such simplifications are often useful, several authors have warned of the risks to \emph{blackboxing}, or ignoring, cognitive mechanisms \cite[p. 658]{Heyes2016, Kendal2018}, given such details may be theoretically meaningful. 
In order to understand how social learning evolves as an adaptation to uncertainty, we 
endowed agents with a biologically plausible cognition capable of integrating information from various sources to maximize payoffs, using \emph{softmax
search}~\cite{Gershman2019}, described in more detail below. 
% \cm{I don't
%   understand this next sentence, but think it can be cut with no loss of meaning:
%   In our model we do not impose any structure in randomness between behaviors, so
% our cognitive model is only sensitive to total randomness.} 
% To do so, we let
% agents see the payoffs to behaviors and use the softmax function to convert these
% observations to a probability distribution from which they strategically select
% their own behavior. %\cm{does this still make sense? and does it need a citation?}

@@

%To understand how uncertainty affects social learning, we developed an agent-based model of social learning evolution.  Evolutionary agent-based models simulate changes in the prevalence of a phenotype over several generations---in this model the phenotype is whether or not an agent engages in social learning.  
In our model we allow just one trait to evolve; social learning, while all other parameters are exogenous. Our primary outcome measure is the overall frequency of the social learning trait at the end of each simulation, and we study how this prevalence responds to the four varieties of uncertainty considered in this paper.

A population of $N$ individuals each must decide which of $B$ behaviors to perform at
each time step within a generation consisting of $L$ time steps. The behavior set is a multi-armed bandit with $B$ arms, that is, agents can pick one of $B$ behaviors with random-valued payoffs~\cite{SuttonBartoBook,McElreath2005,Steyvers2009,Rendell2010,Schulz2019}. \ps{To clarify terminology, the entire set of behaviors is one multi-armed bandit with $B$ arms -- each behavior is not itself a bandit.} 
%Each behavior yields a payoff of either 1 or 0 with some probability that is held fixed within a generation but can change between generations. This probability is therefore equal to the behavior's expected payoff within a generation. 
In each generation, exactly one behavior is more likely to pay off than all the rest and is therefore optimal. Agents
optimize their net payoffs over their lifespan by quickly figuring out which
behavior is optimal and performing it as often as possible within their lifespan.
%The optimal behavior is static within agent lifespans, but varies in general between generations, representing environmental variability.
At the end of each
generation, agents are selected with replacement to reproduce a new generation of $N$ agents, with the probability of reproduction biased by net payoffs. Agents who inherit the social learning trait learn about payoffs from the previous generation, while asocial learners begin life with no \emph{a priori}
knowledge of the world. \cm{or would you consider having flat priors over all behaviors knowledge?}  
%The specifics of the model construction and dynamics are described in detail below.
\ps{I cut a bunch of stuff that came right here as being redundant or unnecessary.}
